Optimal (Multiway) Spatial Joins

### 2024-06-01 msj.tex
yufei

\def\vgap{\vspace{1mm}}

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of intervals in $\real$ using $O(|S|)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log |S| + K)$ time, where $K$ is the number of intervals reported. It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log |S|)$ time. Each interval insertion or deletion on $S$ can be supported in $O(\log |S|)$ time.

Let us first consider the detection version of Problem $\mathscr{A}$. Each point $p = (x, y)$ of $P$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$. We start by sorting the set of y-coordinates defined by all the points and rectangles.

Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in an interval tree $\T$. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. When $\ell$ hits a point $p = (x, y)$ of $P$, search $\T$ to determine if any interval in $S_\ell$ contains the value $x$. If so, point $p$ is covered by at least one rectangle in $R$; otherwise, it is not. The overall running time is $O(n \log n)$.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each vertical segment $v = x \times [y_1, y_2]$ of $V$ is said to {\em define} two y-coordinates $y_1$ and $y_2$. We start by sorting the set of y-coordinates defined by all the horizontal and vertical segments.

Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $V_\ell$ of segments in $V$ that intersect with $\ell$. Let $S_\ell$ be the set of x-coordinates of the segments in $V_\ell$; we store $S_\ell$ in a binary search tree (BST) $\T$. Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ amortized time. When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$. If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$. The overall running time is $O(n \log n)$.

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.}

For Problem $\mathscr{C}$, each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$. We start by sorting the set of y-coordinates defined by all the horizontal segments and rectangles.

Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a structure $\T$ of \cite{aak+12}. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ with {\em weight} $x_2$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $P$, search $\T$ to determine the maximum weight $w$ of all the intervals in $S_\ell$ containing the value $x_1$. If $w \ne -\infty$, we output $(h, p)$ where the point $p$ is defined in a way depending on $w$: if $w \le x_2$, then $p = (w, y)$; otherwise $p = (x_2, y)$. The overall running time is $O(n \log n)$.

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.}

Let us now return to the context of Problem $\mathscr{D}$. Let us first consider the detection version. Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$. We start by sorting the set of y-coordinates defined by all the horizontal segments and rectangles.

We assume, w.l.o.g., that each rectangle of $R$ is given a distinct ID in $[n]$. Initialize an array $A$ of size $|R|$, where cell $A[i]$, $i \in [|R|]$, is allocated to the rectangle $r$ with ID $i$ and will be used to store a constant amount of information for $r$. In the outset, all rectangles of $R$ are marked as inactive (in $A$). During our algorithm, the status of each rectangle will turn from inactive to active at some point, turn from active back to inactive at a later point, and then stay that way forever.

@@

When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a containment query on $\T$ to find all the intervals in $S_\ell$ that are contained by $[z_1, z_2]$; if $K_h$ is the number of such intervals, this retrieval takes $O(\log n + K_h)$ time. For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$). Observe that $h$ is the lowest segment in $H$ that crosses $r$. We store $h$ in the cell of array $A$ allocated to $r$, after which $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time. As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.

@@

Next, we consider the find-all-sorted version of Problem $\mathscr{D}$. Again, initialize an array $A$ of size $|R|$, where each rectangle of $R$ is allocated a cell. For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates. A pointer referencing the linked list is stored in the cell of $A$ allocated to $r$. In the outset, all linked lists are empty. Unlike the detection version, we will not need to keep the active status for the rectangles.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.}

@@

Let us now return to the context of Problem $\mathscr{E}$. We sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$). Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position in the sorted list. We create an array $A$ and assign the $i$-th ($i \in [R]$) cell of $A$ to the rectangle $r \in R$ with ID $i$.

Next, we aim to compute, for each pair $(h, r) \in H \times r$, we output a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$. These pairs may be output in an arbitrary order.

Each horizontal segment $h = [x_1, x_2] \times y$ of $H$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$ is said to {\em define} two y-coordinates $y_1$ and $y_2$. We sort the set of y-coordinates defined by all the horizontal segments and rectangles. Then, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, we maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. We call the rectangle $r$ the {\em host} of $[x_1, x_2]$ and store its ID together with $[x_1, x_2]$ in $\T$. When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a reverse-containment query on $\T$ to find all the intervals in $S_\ell$ that contain $[z_1, z_2]$; the query cost is $O(\log n + K_h)$, where $K_h$ is the number of intervals reported. For each retrieved interval $[x_1, x_2]$, we also obtain the ID $\lambda$ of its host rectangle $r$, and output the pair $(h, \lambda)$.

>>>>

\def\vgap{\vspace{2mm}}
\def\extraspacing{\vspace{2mm}}

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} Consider first the detection version. Sort the set of y-coordinates defined by the points of $P$ and the rectangles of $R$. Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in an interval tree $\T$. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. When $\ell$ hits a point $p = (x, y)$ of $P$, search $\T$ to determine if any interval in $S_\ell$ contains the value $x$. If so, point $p$ is covered by at least one rectangle in $R$; otherwise, it is not. The overall running time is $O(n \log n)$.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$.} Sort the set of y-coordinates defined by all the segments of $H$ and $V$. Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $V_\ell$ of segments in $V$ that intersect with $\ell$. Let $S_\ell$ be the set of x-coordinates of the segments in $V_\ell$; we store $S_\ell$ in a binary search tree (BST) $\T$. Specifically, when $\ell$ hits the lower  (resp., upper) endpoint of a vertical segment $v = x \times [y_1, y_2]$ of $V$, we insert (resp., delete) the value $x$ into (resp., from) $\T$, which can be done in $O(\log n)$ amortized time. When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $H$, search $\T$ to determine if the successor $x'$ of $x_1$ in $S_\ell$. If $x' \le x_2$, then we output a pair $(h, p)$, where $p$ is the point $(x', y)$. The overall running time is $O(n \log n)$.

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$.} Sort the set of y-coordinates defined by all the segments in $H$ and all the rectangles in $R$. Next, sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a structure $\T$ of \cite{aak+12}. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ with {\em weight} $x_2$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. When $\ell$ hits a horizontal segment $h = [x_1, x_2] \times y$ of $P$, search $\T$ to determine the maximum weight $w$ of all the intervals in $S_\ell$ containing the value $x_1$. If $w \ne -\infty$, we output $(h, p)$ where the point $p$ is defined in a way depending on $w$: if $w \le x_2$, then $p = (w, y)$; otherwise $p = (x_2, y)$. The overall running time is $O(n \log n)$.

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$.} Consider first the find-lowest version. Sort the set of y-coordinates defined by all the horizontal segments and rectangles. In the outset, all rectangles of $R$ are marked as inactive. During our algorithm, the status of each rectangle will turn from inactive to active at some point, turn from active back to inactive at a later point, and then stay that way forever.

%We assume, w.l.o.g., that each rectangle of $R$ is given a distinct ID in $[n]$. Initialize an array $A$ of size $|R|$, where cell $A[i]$, $i \in [|R|]$, is allocated to the rectangle $r$ with ID $i$ and will be used to store a constant amount of information for $r$.

@@

When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a containment query on $\T$ to find all the intervals in $S_\ell$ that are contained by $[z_1, z_2]$; if $K_h$ is the number of such intervals, this retrieval takes $O(\log n + K_h)$ time. For each retrieved interval $[x_1, x_2]$, we also obtain its host rectangle $r$ (stored along with $[x_1, x_2]$ in $\T$). Observe that $h$ is the lowest segment in $H$ that crosses $r$; we therefore output $(r, h)$. After that, $r$ is marked as inactive, and accordingly, its x-range $[x_1, x_2]$ is deleted from $T$ in $O(\log n)$ time. As $r$ will remain inactive in the rest of the execution, its x-range will not be retrieved again by another containment query in the future.

@@

Next, we consider the find-all-sorted version of Problem $\mathscr{D}$. Again, initialize an array $A$ of size $|R|$, where each rectangle of $R$ is allocated a cell. For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates. In the outset, all linked lists are empty. Unlike the detection version, we will not need to keep the active status for the rectangles.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$). Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position in the sorted list.

%We create an array $A$ and assign the $i$-th ($i \in [R]$) cell of $A$ to the rectangle $r \in R$ with ID $i$.

@@

Next, we aim to produce, for each pair $(h, r) \in H \times r$, a pair $(h, \lambda)$ where $\lambda$ is the ID of $r$. These pairs may be output in an arbitrary order. Sort the set of y-coordinates defined by all the horizontal segments and rectangles. Sweep (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, we maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in a PST $\T$. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. We call the rectangle $r$ the {\em host} of $[x_1, x_2]$ and store its ID together with $[x_1, x_2]$ in $\T$. When $\ell$ hits a horizontal segment $h = [z_1, z_2] \times y$ of $P$, perform a reverse-containment query on $\T$ to find all the intervals in $S_\ell$ that contain $[z_1, z_2]$; the query cost is $O(\log n + K_h)$, where $K_h$ is the number of intervals reported. For each retrieved interval $[x_1, x_2]$, we also obtain the ID $\lambda$ of its host rectangle $r$, and output the pair $(h, \lambda)$.
